---
output: rmarkdown::github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment="")
devtools::load_all(".")
library(igraph)
library(hypeR)
```

# hypeR Network Extension

[![hypeR](https://img.shields.io/badge/montilab-hypeR-0770d9?labelColor=000000)](https://github.com/montilab/hypeR)
[![](https://lifecycle.r-lib.org/articles/figures/lifecycle-experimental.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![](https://img.shields.io/github/last-commit/montilab/hypeR-next.svg)](https://github.com/montilab/hypeR-next/commits/master)

An extension of hypeR for network enrichment through random walks.

### Installation

```{r, eval=FALSE}
devtools::install_github("montilab/hypeR-next")
```

```{r, eval=FALSE}
library(hypeR.next)
```

### Data

```{r}
data(ig)
```

This package works with undirected `igraph` objects or extensions of `igraph` objects.

```{r}
is(ig)
```

The following vertex attributes are expected at a minimum.

- `name` - Unique node labels
- `symbol` - Node symbols corresponding to the genesets you use  
- `community`- Pre-detected community labels  
- `color` - Community-specific colors  

```{r}
head(igraph::as_data_frame(ig, what="vertices"))
```

```{r, fig.width=12, fig.height=6}
set.seed(1)
layout <- igraph::layout_with_graphopt(ig, 
                                       start=NULL, 
                                       niter=1000,
                                       charge=0.005,
                                       mass=30, 
                                       spring.length=0,
                                       spring.constant=1, 
                                       max.sa.movement=5)

par(mar=c(0,0,0,0))
plot(ig,
     vertex.size=4,
     vertex.color="grey",
     vertex.frame.color="black",
     vertex.label=NA,
     edge.width=1,
     layout=layout)
```

### Network characterization through random walks

```{r}
library(hypeR)
genesets <- hypeR::msigdb_gsets("Homo sapiens", "H", clean=TRUE)
```

```{r}
ig.e <- enrich_communities_static(ig, genesets, restart=0.5, fdr=0.01, top=1)
```

```{r, fig.width=12, fig.height=8}
par(mar=c(0,0,0,0))
set.seed(1234)
plot(ig.e,
     vertex.size=5,
     vertex.label=V(ig.e)$enrichment,
     vertex.color=adjustcolor(V(ig.e)$color, alpha.f=0.6),
     vertex.frame.color=adjustcolor("#000000", alpha.f=0),
     vertex.label.family="Helvetica",
     vertex.label.color="black",
     vertex.label.font=2,
     vertex.label.cex=0.8,
     vertex.label.dist=runif(vcount(ig.e), -0.25, 0.25),
     edge.color=adjustcolor("#000000", alpha.f=0.2),
     layout=layout)
```

Please read the documentation for a comprehensive overview of functionality.
